A Dearth of Young Stars and Planets in the Kepler Field

### 2024-04-25 paper/ms.tex
update ms

%\title{Gyrochrone and Lithium Ages for Stars and Planets Observed by Kepler}
%Ages of the Kepler Stars and Planets
\title{Ages of the Kepler Stars and Planets}

@@

% <250 words
\begin{abstract}
  Recent studies of stars in open clusters have shown how the rotation
  rates and lithium content of FGK stars evolve during the first few
  billion years of their lives.
  %
  Here we apply this knowledge to planets and stars observed by the
  primary Kepler mission, and calculate the implied rotation and
  lithium-based ages.
  %
  Rotation periods are drawn from previous Kepler analyses; lithium is
  measured from new and archival HIRES spectra.
  %
  We report rotation-based ages for \nuniqstarsantosrotgyroappl\ stars
  and \nplwgyroagenograzing\ planets for which we believe
  gyrochronology is probably valid.
  %
  The independent lithium ages show X,Y,Z, though they are subject to their own systematic issues.
  %
  The relative statistical precisions from rotation, $\sigma_{t_{\rm
  gyro}} / t_{\rm gyro}$, are XX +YY -ZZ\%, which is X.X +Y.Y -Z.Z more
  statistically precise than isochronal ages for the same stars, and
  $\approx$X.X times less precise than cluster-based ages.
  %
  The results include \nplyounggyrotwosigma\
  planets younger than 1\,Gyr at 2$\sigma$ (\nplyounggyrothreesigma\
  planets at 3$\sigma$).
  %
  While lithium provides only marginal added information for most of
  this sample,
  we report new robust ages from both rotation and lithium for 14
  confirmed $<$500\,Myr planets, including five mini-Neptunes, five
  super-Earths, three Earths and sub-Earths, and a Neptune-sized
  planet.
  %
  In the broader Kepler target star sample, we find that the
  gyrochrone age distribution rises linearly from 0 to 3\,Gyr, with
  stars 2\,Gyr old being twice as common in the Kepler field as stars
  1\,Gyr old.
  %
  This parent stellar age distribution bounds the relative balance of
  young and old planets detectable not only by Kepler, but also by
  future wide-field transit surveys.
  %
%  The sample of detected Kepler planets has a similar age
%  distribution, though it is slightly deficient at $\lesssim$0.2\,Gyr,
%  likely due to decreased detection sensitivity.
  %
  %Combining our stellar age catalog with Kepler's completeness products, we perform
  %a hierarchical bayesian analysis to infer the average number of planets per star
  %as a function of both stellar and planetary parameters.
  %
  %While we recover many previously known correlations with stellar mass and metallicity,
  %our analysis also highlights the importance of age in shaping the ``sub-Saturn desert'',
  %and in influencing the evolution of the mini-Neptune and super-Earth populations.
  %TODO will this be the case?
%  We also comment on a few previously reported evolutionary trends in
%  the size evolution of close-in mini-Neptunes (2-4\,\rearth) and
%  super-Earths (1-2\,\rearth), a few of which we recover at comparable
%  ($\sim$3$\sigma$) significance to previous work, and one of which we
%  note for the first time.
\end{abstract}

@@

Exoplanet science is in an age of abundance, fueled by the discovery
of thousands of worlds orbiting close to their host stars (CITE, CITE,
CITE).
However, most known exoplanets have ages comparable to that of the solar
system.
This unfortunate fact leaves many gaps in our knowledge of
exoplanetary origins, particularly regarding to degree to which
dynamical migration (CITE), thermal cooling (CITE), and atmospheric
loss (CITE) sculpt planetary systems after the disk disperses.
Finding and characterizing young ($<$1\,Gyr) exoplanets offers one
possible workaround, since these objects can show both individual (CITE,
CITE) and population-level (CITE) evidence for these types of
evolutionary processes.

@@

%While the youngest 0.1\% of stars ($\lesssim$10\,million
%years; Myr) are perhaps the most constraining for studies of planet
%formation, they are extremely challenging targets for close-in
%exoplanet detection 
%\citep[see e.g.][]{Damasso_2020,Bouma_2020_ptfo,Donati_2020}.  Studies of
%the early orbital, structural, and atmospheric evolution of exoplanets
%are perhaps most technically feasible for the 
%$\approx$1\% of stars
%with ages between 10 and 100\,Myr, or perhaps even the 10\% of stars
%younger than 1\,Gyr.  However, since stellar ages are roughly
%uniformly distributed in the solar neighborhood, there is a limit on
%the sample size of young planets that will be detectable in these age
%regimes, particularly relative to their older breathren.

@@

%\begin{figure}[!t]
%	\begin{center}
%		\leavevmode
%		\includegraphics[width=0.49\textwidth]{f1.pdf}
%	\end{center}
%	\vspace{-0.6cm}
%	\caption{
%		{\bf Radii, orbital periods, and ages of transiting exoplanets}.
%    Planets younger than 1\,Gyr with ages more precise than a
%    factor of three are emphasized. 
%    ({\bf cite relevant papers?})
%    Parameters are from the \citet{PSCompPars}.
%		\label{fig:rp_period_age}
%	}
%\end{figure}

@@

@@

To calculate lithium ages, we use \texttt{EAGLES} (git commit
\texttt{ac09637}), which uses a similar method as
\texttt{gyro-interp}, and was calibrated on stars observed by the
Gaia-ESO spectroscopic survey in 52 open clusters with ages spanning 2
to 6{,}000\,Myr \citep{Jeffries_2023}.  We adopt a linear prior in
age, spanning $10^6$ to $10^{10.1}$\,yr, and symmetrize our EW
uncertainties for purposes of interfacing with \texttt{EAGLES} by
taking the maximum of the measured positive and negative 1$\sigma$
uncertainty intervals.
Certain quality flags described in Section~\ref{subsec:flags} are also
relevant for lithium ages, specifically those associated with
crowding and binarity since those effects can produce both systematic
and astrophysical biases in EW measurements.

>>>>

%\title{Gyrochrone and Lithium Ages for Stars and Planets Observed by Kepler}
%Ages of the Kepler Stars and Planets
\title{A Dearth of Young Stars and Planets in the Kepler Field}

@@

% <250 words
\begin{abstract}
  Recent studies of stars in open clusters have shown how the rotation
  rates and lithium content of FGK stars evolve during the first few
  billion years of their lives.
  %
  Here we apply this knowledge to planets and stars observed by the
  primary Kepler mission, and calculate the implied rotation and
  lithium-based ages.
  %
  Rotation periods are drawn from previous Kepler analyses; lithium is
  measured from new and archival HIRES spectra.
  %
  We report rotation-based ages for \nuniqstarsantosrotgyroappl\ stars
  and \nplwgyroagenograzing\ planets for which we believe
  gyrochronology is probably valid.
  %
  The independent lithium ages show X,Y,Z, though they are subject to
  their own systematic issues.
  %
  The relative statistical precisions from rotation, $\sigma_{t_{\rm
  gyro}} / t_{\rm gyro}$, are XX +YY -ZZ\%, which is X.X +Y.Y -Z.Z more
  statistically precise than isochronal ages for the same stars, and
  $\approx$X.X times less precise than cluster-based ages.
  %
  The results include \nplyounggyrotwosigma\ planets younger than
  1\,Gyr at 2$\sigma$ (\nplyounggyrothreesigma\ planets at 3$\sigma$).
  %
  While lithium provides only marginal added information for most of
  this sample, we report new robust ages from both rotation and
  lithium for 14 confirmed $<$500\,Myr planets, including five
  mini-Neptunes, five super-Earths, three Earths and sub-Earths, and a
  Neptune-sized planet.
  %
  In the broader Kepler target star sample, we find that the
  gyrochrone age distribution rises linearly from 0 to 3\,Gyr, with
  stars 2\,Gyr old being twice as common in the Kepler field as stars
  1\,Gyr old.
  %
  This parent stellar age distribution bounds the relative balance of
  young and old planets detectable not only by Kepler, but also by
  future wide-field transit surveys.
  %
\end{abstract}

@@

Exoplanet science is in an age of abundance, fueled by the discovery
of thousands of worlds orbiting close to their host stars
\citep{Borucki10,2015JATIS...1a4003R}.
Most of these exoplanets have ages comparable to that of the solar
system.
This fact leaves many gaps in our knowledge of what exact dynamical
and physical processes produce these objects.
Finding and characterizing young ($\lesssim$1\,Gyr) exoplanets offers
one possible workaround.
The reason is that processes such as thermal cooling, atmospheric
loss, and dynamical instability occur over $\approx$10\,Myr to
$\approx$1\,Gyr timescales in a way that can leave observable imprints in
the properties of young planets (CITE theory reviews).
Individual exemplars include V1298~Tau \citep{David_2019}, an
adolescent resonant chain of four transiting puffy planets,
and HIP~67522 \citep{Rizzuto_2020}, a Jupiter-sized planet with a
Neptunian mass (CITE).
Population-level examples include tentative differences in the size
distribution of young vs.~old planets
(CITEVach,CITEChristianse,CiteDavid).

@@

@@

@@

%     While the youngest 0.1\% of stars ($\lesssim$10\,million
%     years; Myr) are perhaps the most constraining for studies of planet
%     formation, they are extremely challenging targets for close-in
%     exoplanet detection 
%     \citep[see e.g.][]{Damasso_2020,Bouma_2020_ptfo,Donati_2020}.  Studies of
%     the early orbital, structural, and atmospheric evolution of exoplanets
%     are perhaps most technically feasible for the 
%     $\approx$1\% of stars
%     with ages between 10 and 100\,Myr, or perhaps even the 10\% of stars
%     younger than 1\,Gyr.  However, since stellar ages are roughly
%     uniformly distributed in the solar neighborhood, there is a limit on
%     the sample size of young planets that will be detectable in these age
%     regimes, particularly relative to their older breathren.

@@

To calculate lithium ages, we use \texttt{EAGLES} (git commit
\texttt{ac09637}), which uses a similar method as
\texttt{gyro-interp}, and was calibrated on stars observed by the
Gaia-ESO spectroscopic survey in 52 open clusters with ages spanning 2
to 6{,}000\,Myr \citep{Jeffries_2023}.  We adopt a linear prior in
age, spanning $10^6$ to $10^{10}$\,yr, and symmetrize our EW
uncertainties for purposes of interfacing with \texttt{EAGLES} by
taking the maximum of the measured positive and negative 1$\sigma$
uncertainty intervals.
Certain quality flags described in Section~\ref{subsec:flags} are also
relevant for lithium ages, specifically those associated with
crowding and binarity.
For instance, rapid rotators are more common in binaries, and they also
generally have higher lithium abundances than slow-rotating stars of the
same mass (CITE).
Stars for which rotation-based age dating is dubious heavily overlap
with those for which lithium-based ages are dubious.

In part because many of our EWs are upper limits, 
many of the lithium ages are lower limits.
In such cases,  the inferred age posterior is strongly influenced by
{\it i)} the assumed intrinsic dispersion in the lithium model,
and {\it ii)} by the prior.
For instance, 
 we quote the 3$\sigma$ (99.7$^{\rm th}$ percentile) limits.
